ML-based approach for accelerating global search algorithm for solving multicriteria problems\thanks{Supported by organization x.}

### 2022-02-07 2022/LION/main.tex
fix some typos

In the framework of the proposed approach the information-statistical theory of global search is used for solving the scalar optimization problems (\ref{eq:4}). For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000}. While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain D and computation of the convolution (\ref{eq:4}) at this point.

@@

\begin{enumerate}
  \item Obtain univariate images of all points of preceding trials mapped onto one-dimensional interval [0,1] by means of Peano curves.
  \item Partition the segment [0,1] into subintervals according to the images of trials performed.
  \item For each subinterval compute a numerical value called \textit{characteristic} of this subinterval.
  \item Select the subinterval with the highest characteristic.
  \item Place the point of new trial within the subinterval with the highest characteristic. 
  \item Map the new trial point to the point $y \in D$ using Peano evolvent and compute the value of the function (\ref{eq:4}) at this point. 
\end{enumerate}

@@

The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal [0,1].

Let $R(i)$ be the characteristic of the $i$-th subinterval. This characteristic is supposed to consist of two parts:

\begin{equation}
R(i)=R_{ags} (i)+ \alpha R_{PS} (i).
\label{eq:5}
\end{equation}

@@

In the course of Pareto set assessment all trial points are partitioned into two classes. This bipartition is based on belonging to the Pareto set. For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}. Examples of separated planes are shown in Fig. \ref{fig:1}. During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.

@@

\begin{equation}
f_1 (y) = (y_1-1) y_2^2+1, f_2 (y) = y_2, 0\leq y_1, y_2 \leq 1.
\label{eq:6}
\end{equation}

@@

\begin{table}[]
\caption{Efficiency of the multicriterial algorithms compared}
\label{tab:1}
\begin{tabular}{ccccccc}
\hline
Algorithm                      & MC    & SEMO  & NUC   & BLO   & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 0$)\end{tabular} & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 1.5$)\end{tabular} \\ \hline
Number of trials               & 500   & 500   & 515   & 498   & 302             & 269               \\
Number of Pareto points found  & 67    & 104   & 29    & 68    & 92              & 79                \\
HV index (the more the better) & 0.300 & 0.312 & 0.306 & 0.308 & 0.312           & 0.312             \\
DU index (the less the better) & 1.277 & 1.116 & 0.210 & 0.175 & 0.101           & 0.103             \\ \hline
\end{tabular}
\end{table}

@@

\begin{equation}
\begin{matrix}
f(y) = - (AB + CD)^{1/2} \\
AB = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [A_{ij} a_{ij}(y_1, y_2) + B_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\
CD = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [C_{ij} a_{ij}(y_1, y_2) - D_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\ 
a_{ij}(y_1, y_2) = \sin(\pi i y_1) \sin(\pi j y_2), b_{ij}(y_1, y_2) = \cos(\pi i y_1) \cos(\pi j y_2),
\end{matrix}
\end{equation}
$0 \leq y_1,y_2 \leq 1$, have been taken where parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly \cite{Grishagin2015_2,Gergel2019_2}.

@@

\begin{table}[]
\caption{Efficiency of the multicriterial algorithms compared}
\label{tab:2}
\begin{tabular}{ccc}
\hline
                                  & ML\_MGSA ($\alpha=0.0$) & ML\_MGSA ($\alpha=0.01$) \\ \hline
Average number of iterations      & 1902.4           & 658.6             \\
Average value of  DU index        & 1.32             & 1.55              \\
Average value of HV index         & 92.1             & 91.9              \\
Reducing the number of iterations & 1                & 2.9     \\  \hline        
\end{tabular}
\end{table}

>>>>

In the framework of the proposed approach the information-statistical theory of global search is used for solving the scalar optimization problems (\ref{eq:4}). For problems with several variables ($N>1$) the dimensionality reduction on the base of Peano-type mappings \cite{Sergeyev2013,Strongin2000}. While solving a problem (\ref{eq:4}) the algorithm generates a sequence of trials where the term ``trial'' means evaluation of criteria at a point of the feasible domain $D$ and computation of the convolution (\ref{eq:4}) at this point.

@@

\begin{enumerate}
  \item Obtain univariate images of all points of preceding trials mapped onto one-dimensional interval $[0,1]$ by means of Peano curves.
  \item Partition the segment $[0,1]$ into subintervals according to the images of trials performed.
  \item For each subinterval compute a numerical value called \textit{characteristic} of this subinterval.
  \item Select the subinterval with the highest characteristic.
  \item Place the point of new trial within the subinterval with the highest characteristic. 
  \item Map the new trial point to the point $y \in D$ using Peano evolvent and compute the value of the function (\ref{eq:4}) at this point. 
\end{enumerate}

@@

The second improvement is connected with a new method for calculating the characteristics of subintervals formed during univariate optimization on the internal $[0,1]$.

Let $R(i)$ be the characteristic of the $i$-th subinterval. This characteristic is supposed to consist of two parts:
\begin{equation}
R(i)=R_{ags} (i)+ \alpha R_{PS} (i).
\label{eq:5}
\end{equation}

@@

In the course of Pareto set assessment all trial points are partitioned into two classes. This bipartition is based on belonging to the Pareto set. For these classes, a separation hyperplane is constructed in the domain of criteria values using the logistic regression \cite{Yu2011}. Examples of separated planes are shown in Fig.~\ref{fig:1}. During the choice of a new trial point, the distances of all obtained criteria values from the separated plane are calculated and the value of $R_{PS}(i)$ is built depending on these distances.

@@

\begin{equation}
f_1 (y) = (y_1-1) y_2^2+1, \; f_2 (y) = y_2, \; 0\leq y_1, y_2 \leq 1.
\label{eq:6}
\end{equation}

@@

\begin{table}
\caption{Efficiency of the multicriterial algorithms compared}
\label{tab:1}
\begin{tabular}{ccccccc}
\hline
Algorithm                      & MC    & SEMO  & NUC   & BLO   & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 0$)\end{tabular} & \begin{tabular}[c]{@{}c@{}}ML\_MGSA \\ ($\alpha= 1.5$)\end{tabular} \\ \hline
Number of trials               & 500   & 500   & 515   & 498   & 302             & 269               \\
Number of Pareto points found  & 67    & 104   & 29    & 68    & 92              & 79                \\
HV index (the more the better) & 0.300 & 0.312 & 0.306 & 0.308 & 0.312           & 0.312             \\
DU index (the less the better) & 1.277 & 1.116 & 0.210 & 0.175 & 0.101           & 0.103             \\ \hline
\end{tabular}
\end{table}

@@

\begin{equation}
\begin{matrix}
f(y) = - (AB + CD)^{1/2} \\
AB = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [A_{ij} a_{ij}(y_1, y_2) + B_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\
CD = \Big(\sum_{i=1}^7 \sum_{j=1}^7 [C_{ij} a_{ij}(y_1, y_2) - D_{ij} b_{ij}(y_1, y_2)]\Big)^2 \\ 
a_{ij}(y_1, y_2) = \sin(\pi i y_1) \sin(\pi j y_2), b_{ij}(y_1, y_2) = \cos(\pi i y_1) \cos(\pi j y_2),
\end{matrix}
\end{equation}
$0 \leq y_1,y_2 \leq 1$, have been taken where parameters $-1 \leq A_{ij},B_{ij},C_{ij},D_{ij} \leq 1$ are the independent random numbers distributed uniformly \cite{Gergel2019_2,Grishagin2015_2}.

@@

\begin{table}
\caption{Efficiency of the multicriterial algorithms compared}
\label{tab:2}
\begin{tabular}{ccc}
\hline
                                  & ML\_MGSA ($\alpha=0.0$) & ML\_MGSA ($\alpha=0.01$) \\ \hline
Average number of iterations      & 1902.4           & 658.6             \\
Average value of  DU index        & 1.32             & 1.55              \\
Average value of HV index         & 92.1             & 91.9              \\
Reducing the number of iterations & 1                & 2.9     \\  \hline        
\end{tabular}
\end{table}
